A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-08-11 docs/model_JAM_v1.01.tex

[TO DO]

>>>>

Jeans axisymmetric models (JAM) assume galaxies (1a) to be collisionless, i.e. the collisionless Boltzmann equation for the distribution function $f(\vect{x},\vect{v},t)$ has to be satisfied ($\frac{\diff f(\vect{x},\vect{v},t}{\diff t} = 0$), (1b) in a steady state ($\frac{\partial}{\partial t} = 0$), (1c) axisymmetric (best described in cylindrical coordinates $(R,z,\phi)$ and $\frac{\partial}{\partial \phi} = 0$) and (2) to have their velocity ellipsoid aligned with the cylindrical coordinate system. From this follow the Jeans equations as the vector-valued first moment of the Boltzmann equation, i.e.
\begin{equation*}
[TO DO]
\end{equation*}
[TO DO]

### 2015-08-11 docs/model_lensing_v1.01.tex

>>>>

\paragraph{Model fitting.} As described above our lensing model has the following free parameters: the source position $(\xi,\eta)$, and the radial slope $\alpha$ and Fourier parameters $(a_k,b_k)$ of the lens mass distribution in eq. (\ref{eq:scalefreemodel}) and (\ref{eq:Fourieransatz}). We want to find the lensing model which minimizes for all four images the distance between the observed image positions $\vec{\theta}_{oi}$ and those predicted by the lensing model $\vec{\theta}_{pi}$. Because we want to avoid solving the lens equation (cf. eq. (\ref{eq:lenseqpot}) and (\ref{eq:Fourierlenseq})) for $\theta_{pi}$, we follow \citet{1991ApJ...373..354K} and cast the calculation back to the source plane using the magnification tensor in eq. (\ref{eq:magnificationtensor}):
\begin{eqnarray*}
\chi^2_\text{lens} &=& \sum_{i=1}^{4} \left|\left( \begin{matrix} \frac{1}{\Delta_x} & 0\\0 & \frac{1}{\Delta_y}\end{matrix}\right) \left( \vec{\theta}_{pi} - \vec{\theta}_{oi} \right)\right|^2\\
&\simeq& \sum_{i=1}^{N} \left|\left( \begin{matrix} \frac{1}{\Delta_x} & 0\\0 & \frac{1}{\Delta_y}\end{matrix}\right)  \left.\mathscr{M}\right|_{\vec{\theta}=\vec{\theta}_{oi}} \left( \begin{matrix} \xi - \tilde{\xi}_i \\ \eta - \tilde{\eta}_i \end{matrix} \right) \right|^2,
\end{eqnarray*}
where $(\Delta_x,\Delta_y)$ are the measurement errors of the image positions $\vec{\theta}_{oi}$. $\left.\mathscr{M}\right|_{\vec{\theta}=\vec{\theta}_{oi}}$ is the magnification tensor and $(\tilde{\xi}_i,\tilde{\eta}_i)$ the source position according to the lens equation evaluated at $\vec{\theta}_{oi}$. Following \citet{GlennEC} we add a term
\begin{equation*}
\chi^2_\text{shape} = \lambda \sum_{k \geq 3} \frac{\left(a_k^2 +b_k^2 \right)}{a_0^2} 
\end{equation*}
which forces the shape of the mass distribution to be close to an ellipse. The total $\chi^2$ to minimize is therefore
\begin{equation*}
\chi^2 = \chi^2_\text{lens} + \chi^2_\text{shape}
\end{equation*}
We set $a_1 = b_1 = 0$, which corresponds to the choice of origin; in this case the center of the galaxy.
\\To be able to constrain the slope $\alpha$, we would have needed flux ratios for the images as in \citet{GlennEC}. But the extended quality of the images and the uncertainty in surface brightness subtraction makes flux determination too unreliable and we do not include them in the fitting. Even though the constraint from just the image position fit on $\alpha$ is very weak, we were however able to show that the image positions in tab. \ref{tab:lenspos} minimize $\chi^2$ at $\alpha=1$ and also our other image position sets from different models and filters are consistent with a flat rotation curve. In the following we therefore set $\alpha=1$.
